A Practical Guide for Generating Reproducible and Programmatic Neuroimaging Visualizations

### 2023-02-10 manuscript.Rmd
sp fx

The visualization of neuroimaging data is one of the primary ways in which we evaluate data quality, interpret results, and communicate findings. These visualizations are commonly produced using graphical user interface-based (GUI) tools where individual images are opened and, within each instance, display settings are manually changed until the desired output is reached. In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility. However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community. These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code, especially within the R, Python and MATLAB environments. Here, we first discuss the rational for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools. We then provide a practical guide outlining the steps required to make code-based brain visualizations and provide a web-app (<https://sidchop.shinyapps.io/braincode/>; Figure 4) which can generate simple code-templates as starting points for these visualizationds. We also present a comprehensive table of tools currently available for programmatic brain visualizations (Table 1) and provide didactic examples of visualizations and associated code as a point of reference (Figure 2-3). The focus of this guide is on human brain magnetic resonance imaging (MRI) data, but many of the principals discussed and tools provided will equally apply to visualizing data from other organs and imaging modalities such as EEG, MEG, PET and CT.

>>>>

The visualization of neuroimaging data is one of the primary ways in which we evaluate data quality, interpret results, and communicate findings. These visualizations are commonly produced using graphical user interface-based (GUI) tools where individual images are opened and, within each instance, display settings are manually changed until the desired output is reached. In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility. However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community. These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code, especially within the R, Python and MATLAB environments. Here, we first discuss the rationale for the widespread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools. We then provide a practical guide outlining the steps required to make code-based brain visualizations and provide a web-app (<https://sidchop.shinyapps.io/braincode/>; Figure 4) that can generate simple code-templates as starting points for these visualizations. We also present a comprehensive table of tools currently available for programmatic brain visualizations (Table 1) and provide didactic examples of visualizations and associated code as a point of reference (Figure 2-3). The focus of this guide is on human brain magnetic resonance imaging (MRI) data, but many of the principles discussed and tools provided will equally apply to visualizing data from other organs and imaging modalities such as EEG, MEG, PET and CT.
